Jane, John … Leslie? A Historical Method for Algorithmic Gender Prediction

### 2015-08-05 article.Rmd
Who needs MS Word?
Slightly less jargony abstract

---
title: "Jane, John … Leslie? A Historical Method for Algorithmic Gender Prediction"
author: "Cameron Blevins and Lincoln Mullen"
output:
  html_document: default
  word_document: default
---

@@

This paper outlines a new method for inferring the gender of personal names in large, machine-readable datasets. In contrast to existing methods of predicting gender that treat names as static entities, this method uses a humanistic approach to take into account the dynamic nature of naming practices. It uses historical data to measure the likelihood that a name was associated with a particular gender based on the time or place under study. This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a flexible tool to apply and adapt to analyses of gender. The article first describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/). It goes on to apply the method to a case study in which we examine gender and gatekeeping in the American historical profession over the past half-century. The Gender package illustrates the importance of incorporating humanistic approaches to computer science and related fields.

>>>>

---
title: "Jane, John … Leslie? A Historical Method for Algorithmic Gender Prediction"
author: "Cameron Blevins and Lincoln Mullen"
output:
  html_document: default
---

@@

This article describes a new method for inferring the gender of personal names using large historical datasets. In contrast to existing methods of predicting gender that treat names as if they are timelessly associated with gender, this method uses a historical approach that takes into account the dynamic nature of naming practices. It uses historical data to measure the likelihood that a name was associated with a particular gender based on the time or place under study. This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a tool to estimate the gender of names. The article first describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/). It goes on to apply the method to a case study in which we examine gender and gatekeeping in the American historical profession over the past half-century. The Gender package illustrates the importance of incorporating historical approaches to computer science and related fields.
